A Possibilistic Valid-time Model

### 2011-12-15 ipmu2012/conclusions.tex
Last update

We presented a valid-time model to represent and querying ill-known temporal intervals. The main advantage of the framework is that it is always possible to get both a possibility and a necessity measures for every comparison, which is also useful for ranking purposes. The framework models the Allen's relations and it has the flexibility to model specific and more complex relations by means of ill-known constraints.  As future work, the time interval in the query specification would also be ill-known.

>>>>

We presented a valid-time model to represent and query ill-known temporal intervals. The main advantage of this framework is that it works with ill-known intervals and thus tries to correctly represents and handle querying temporal uncertainty in the database by providing both possibility and necessity measures. 
%The final ranking step provides a useful result record set. 
It is possible to model the Allen's relations and also more complex relations. As future work, allowing the time interval in the query specification to be ill-known and new ways for aggregation and ranking are considered.

### 2011-12-15 ipmu2012/introduction.tex
Last update

Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.

First of all, the concept of \emph{time} itself is very complex. Klein~\cite{klein94} studied the concept of time in language and Devos~\cite{devos94} modelled vague temporal expressions by means of fuzzy sets~\cite{zadeh65}. More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}. Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.

@@

To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy,Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04,schockaert08}.

Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen91,Snodgrass84,Nascimento95} that consider time as a crisp notion. In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97,Garrido2009}.

@@

In this work we will propose a possibilistic valid time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.

>>>>

Dealing with time in information systems is usually a difficult task~\cite{Bolour82}.
First of all, the concept of \emph{time} itself is very complex. More towards information systems, several proposals have been concerned with the obtaining of theoretical models that allow the representation of time in information systems~\cite{Cru97}. Notably, temporal relations between time intervals and to a lesser extent time points, were first studied by Allen~\cite{Allen83}.

@@

To allow information systems to cope with these imperfections, many approaches adopt fuzzy sets for the representation of temporal information~\cite{Billiet:Pons:Matthe:DeTre:Pons:2011:BipolarFuzzy},\cite{Dubois:jucs_9_9:fuzziness_and_uncertainty_in}. The temporal relations studied by Allen were recently fuzzified by several authors~\cite{ohlbach04},\cite{schockaert08}.

Next to the issues in temporal modelling, the addition of time to a relational database leads to several more practical problems, as the modelling of a time-variant object has an impact on the database consistency. Different models have been proposed \cite{Jensen91},\cite{Snodgrass84},\cite{Nascimento95} that consider time as a crisp notion. In recent works, some authors consider the necessity of allowing imprecision in temporal models~\cite{Cru97},\cite{Garrido2009}.

@@

In this work we will propose a possibilistic valid-time model for relational databases, based on the framework presented in~\cite{Pon11}, which deals with the inference of uncertainty about the evaluation of sets from the uncertainty about the exact values used in evaluation.

### 2011-12-15 ipmu2012/ipmu.tex
Last update

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{multirow}
\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}

@@

%
\section*{Acknowledgements}
%
The researchers are supported by the grant BES-2009-013805 within the research project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in relational DBMS}.

>>>>

\usepackage[dvips]{graphicx}
\usepackage{makeidx}  % allows for indexgeneration
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{multirow}

\hyphenation{pro-per-ties}
\hyphenation{ge-ne-ral-ly}
\hyphenation{pre-fe-ren-ces}
\hyphenation{u-sing}
\hyphenation{pu-nish-ment}

@@

%
\section*{Acknowledgements}
%
The researchers are supported by the grant BES-2009-013805 in the project TIN2008-02066: \emph{Fuzzy Temporal Information treatment in RDBMS}.

### 2011-12-15 ipmu2012/preliminaries.tex
Last update

To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set satisfying both constraints is then:

@@

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A \emph{temporal database} is a database that manages some aspects of time in its schema. The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior. A \emph{chronon} is the shortest duration of time supported by the database. Time can be represented either as points or intervals \cite{655777}.
%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.

@@

\begin{itemize}
	\item
	\emph{Transaction time}: The time when the fact is stored in the database.
	\item
	\emph{Valid time}: The time when the fact is true in the modelled reality.
	\item
	\emph{Decision time}: The time when an event was decided to happen. 
\end{itemize}

Database models can then be classified into transaction time \cite{Jensen91}, decision time, valid time, bi-temporal \cite{Snodgrass84}(both valid and transaction time) or tri-temporal \cite{Nascimento95} (valid, transaction and decision time) models.

To deal with time points \cite{Dubois89} or intervals \cite{Garrido2009} that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.

>>>>

To calculate possibility and necessity concerning a conjunction of constraints, the $\min$ operator is used. The possibility and necessity of the set is: %satisfying both constraints is then:

\vspace{-10pt}

@@

\subsection{Temporal Databases}
%In this section, the proposed reasoning is applied to the specific context of intervals on the real line. This setting is of specific interest in the context of fuzzy temporal databases.
A \emph{temporal database} is a database that manages some aspects of time in its schema \cite{Dyreson1994}. The reality a temporal database tries to model, contains some temporal notions which have to be handled specifically in order to maintain a consistent modelling behavior. A \emph{chronon} is the shortest duration of time supported by the database. Time can be represented either as points~\cite{Dubois89} or intervals ~\cite{Garrido2009},\cite{655777} that may be subject to imperfection.
%There are proposals for the fuzzyfication of the time point and the fuzzyfication of the time interval.

@@

\begin{itemize}
	\item
	\emph{Transaction time}~\cite{Jensen91}: The time when the fact is stored in the database.
	\item
	\emph{Valid time}~\cite{Snodgrass84}: The time when the fact is true in the modelled reality.
	\item
	\emph{Decision time}~\cite{Nascimento95}: The time when an event was decided to happen. 
\end{itemize}

Database models can also be classified into \emph{bi-temporal} (both valid and transaction-time) or \emph{tri-temporal}  (bi-temporal and decision time) models.

%To deal with time points  or intervals  that are subject to imperfection and can thus not be precisely known, fuzzy temporal models \cite{schockaert08} exist. %Some fuzzy temporal models assume that the time stored in the database is an interval. The temporal interval is represented by two ill-known time points: $X$  an ill-known starting point and $Y$ an ill-known ending point. The interval $\left[X,Y\right]$ is not a fuzzy interval but an ill-known interval: it is a crisp interval but it is partially unknown which values are in this interval.
